% !TeX -lualatex 
\documentclass[ba,logo]{ivs-thesis} 

\input{preamble} 

\title{Unpaired Image Translation for Near-Infrared Image Colorization} 
\author{Ayk Borstelmann\\\birthinfo{March 28, 2002 in Aschaffenburg}\\\matricnumber{3441004}}

\examiner{%
	PD Dr. Volker Steinhage\\University of Bonn \and 
	Prof. Dr. Matthias B. Hullin\\University of Bonn
}
\supervisor{Timm Haucke\\ University of Bonn}

\date{September 18, 2023}

\addbibresource{references.bib} 
\usepackage{blindtext}


\begin{document}
\frontmatter

\maketitle

\cleardoublepage
\begin{abstract}
	\todo{Refine}
	Unpaired image translation problems belong to the broader category of image-to-image translation, 
	wherein the goal is to learn a translation without any available image pairs. 
	On instance of this problem is Near-Infrared colorization, which aims to convert NIR images to colored RGB images,
	thereby benefiting from both formats:
	While NIR images are suitable for capturing images in low-light conditions without visible flash, 
	RGB images facilitate human comprehension due to their ability to discriminate colors. 
	Current methods for this problem involve Generative Adversial Networks (GANs), 
	which may suffer from mode collapse and unrealistic images outcomes. 
	Diffusion models are known for their extremely realistic image synthesis. 
	We study unpaired image translation on such models. 
	This tooling can benefit researchers working with camera traps using NIR images as well as object detection systems.
\end{abstract}

\chapter*{Acknowledgements}

\cleardoublepage
\chapter*{Statement of Authorship}
\input{authorship-en.tex}
%\chapter*{Eigenst\"{a}ndigkeitserkl\"{a}rung}
%\input{authorship-de.tex} 

\cleardoublepage
\tableofcontents


\mainmatter

\include{introduction}

\include{methods}

\include{evaluations}

\include{conclusions}

% Workaround: 
% The backmatter deactivates chapter numbering for the appendices which are 
% supposed to appear in the backmatter _after_ the bibliography. This makes 
% referencing the appendices a hassle. A workaround is to put the
% backmatter, bibilography, and lists-of in a LaTeX group using \begingroup 
\begingroup

\backmatter

\printbibliography[heading=bibintoc]

\listoffigures

\listoftables

\endgroup
% End of Workaround 

\appendix

\end{document}